Long-Term Reproducibility through Replication:\\ the ReScience Initiative

### 2017-06-09 draft.tex
Revised and restructured the former "Discussions" section

\section*{Discussion}

% -> Short-term vs long-term reproducibility

There are several different reasons for non-reproducibility. For the most part,
those reasons are specific to each scientifc domain but, from our publishing
experience, there are also some common behaviors that can be identified.
Missing code and/or data, unknown dependencies, inaccurate or imprecise
description appears to be shared characteristic of a lot of non-reproducible
work. If the code accompanying a ReScience replication can fix a lot of these
issue, it can nonetheless only ensure short-term reproducibility. Even if a
piece of code is written following best practices, and then reviewed and
tested, it suffers from the same problem as any other piece of code, i.e. it
depends on a software stack whose stability is not guaranteed in the long
term. \citep{Mesnard:2016} illustrated this problem very clearly when they
tried to reproduce their own work performed two years earlier. Even though
Barba's group is committed to reproducible research practices, they did not
escape the many problems one can face when trying to re-run a piece of code.
This means the newly produced code for ReScience will likely be obsolete at
some point in the future. It can be a matter of months or years, or even
decades, but ultimately, obsolescence is inevitable. Therefore, the long-term
value of a ReScience publication is not the actual code but the accompanying
article. If you consider the original article with missing or incomplete
information and the new article with missing and checked information, you have
a complete and consistent description of the original work. 5, 10 or 20 years
from now, anyone should be able to replicate the work thanks to these two
articles. Of course, the new code can also help, but the true value of a
replication is really the accompanying article.\\

One immediate and legitimate question is to wonder to what extent such
replication could be performed prior to the publication of the original
article. This would strongly reinforce a claim because a successful and
independent replication would have been performed prior to the publication. As
illustrated on figure \ref{fig:coscience}, this would require for a group A to
contact a group B and to send them their a draft of their original work (the
one that would be normally submitted to a journal) such that group B could run
a replication and confirm or infirm the work. In case of confirmation, a
certified article could be later published with both group as authors (each
group being identified according to their respective roles). However, if the
replication failed and the original work cannot be fixed, this would prevent
the publication. This model would improve the overall reproducibiloty and would
slow down considerably the pace of publication we're observing
today. Unfortunately, such scenario is highly improbable. The pressure to
publish is so strong and the incentive for doing replication so low that it
would most probably prevent such collaborative work. At this point, CoScience
is mostly an utopia for tomorrow.

>>>>

\section*{Lessons learned}

Although ReScience is still a young project, the submissions handled
so far already provide valuable experience concerning the
reproducibility and replicability of computational work in scientific
research.

\subsection*{Short-term and long-term reproducibility}

While some of the reasons for non-reproducibility are specific to each
scientifc domain, our experience has shown that there are also some
common issued that can be identified.  Missing code and/or data,
undocumented dependencies, and inaccurate or imprecise description
appear to be characteristic of much non-reproducible work. Moreover,
these problems are not always easy to detect even for attentive
reviewers, as we discovered when some articles published in ReScience
turned out to be difficult to reproduce for someone else for exactly
the reasons listed above. ReScience reviewers are scientists working
in the same domain as the submitting authors, because familiarity with
the field is a condition for judging if a replication is
succesful. But this also means that our reviewers share a significant
common background with the authors, and that background includes the
software packages and programming languages adopted by their
community.  In particular, if both authors and reviewers have
essential libraries of their community installed on their computers,
they may not notice that these libraries are actually dependencies of
the submitted code.  While solutions to this problem evidently exist
(ReScience could, for example, request that authors make their
software work on a standard computational environment supplied in the
form of a virtual machine), they represent an additional effort to
authors and therefore discourage from submitting replication work to
ReScience.  Our goal is to develop a set of recommendations for
authors that represent a workable compromise between reproducibility
and implementation effort.\\

Another issue with reproducibility is that with today's computing
technology, long-term reproducibility can only be achieved by imposing
drastic constraints on languages and libraries that are not compatible
with the requirements of research computing. This problem is nicely
illustrated by \citep{Mesnard:2016} whose authors report trying to
reproduce their own work performed two years earlier. Even though
Barba's group is committed to reproducible research practices, they
did not escape the many problems one can face when trying to re-run a
piece of code. As a consequence, the code newly produced for ReScience
will likely cease to be functional at some point in the future.
Therefore, the long-term value of a ReScience publication is not the
actual code but the accompanying article. The combination of the
original article and the replication article provide a complete and
consistent description of the original work, as evidenced by the fact
that replication was possible. Even 5, 10, or 20 years later, a
competent scientist should be able to replicate the work again thanks
to these two articles. Of course, the new code can also help, but the
true long-term value of a replication is the accompanying article.\\

\section*{Outlook}

One immediate and legitimate question is to wonder to what extent a
replication could be performed \textit{prior} to the publication of
the original article. This would strongly reinforce a claim because a
successful and independent replication would be available right from
the start. As illustrated in figure \ref{fig:coscience}, this would
require for a group A to contact a group B and send them a draft of
their original work (the one that would be normally submitted to a
journal) such that group B could perform a replication and confirm or
infirm the results. In case of confirmation, a certified article could
be later published with both groups as authors (each group being
identified according to their respective roles). However, if the
replication fails and the original work cannot be fixed, this would
prevent publication. This model would improve the quality of
computational research and also slow down considerably the rapid pace
of publication we are observing today. Unfortunately, such a scenario
seems highly improbable today. The pressure to publish is so strong
and the incentive for doing replication so low that it would most
probably prevent such collaborative work. At this point, CoScience is
just a utopia for tomorrow.
